---
title: "R Notebook"
output: html_notebook
---

```{r results='hide'}
library(tidyverse)
library(gganimate)
# library(ggthemes)
library(fontregisterer)
library(av)
```

```{r}
# "Creative Copy on AIIT Homecoming Day Matsui Minoru"
input <- "Creative Copy Matsui Minoru" |> 
  str_to_upper()
randomize_N_times <- 100
nwords <- input |>  
  str_split(pattern = "\\s") |> 
  length()
# init out of for
result <- vector(length = randomize_N_times)
input_vec <- input |> 
  str_split(pattern = "\\s") |> 
  unlist()
for(n in 1:randomize_N_times) {
  # initialize
  result_vecs <- vector(length = nwords)
  for(i in 1:length(input_vec)) {
    input_str <- input_vec[i] 
    length_of_str <- nchar(input_str)
    split_letters <-input_str |>
      str_split(pattern = "") |> 
      unlist()
    if (length_of_str > 2) {
    middle_section <- split_letters[2:(length(split_letters) - 1)] |> 
      sample() |> 
      paste(collapse = "")
    } else {middle_section <- ""}
      result_vec <- paste0(split_letters[1], middle_section, split_letters[length(split_letters)])
      # result_vec = "Mrionu"
      result_vecs[i] <- result_vec
  } 
  result_str <- paste(result_vecs, collapse = " ")
  result[n] <- result_str
}
```

## copypaste
```{r}
result |> paste(collapse = "lalala")
```



# grouped

now group some letters into chunks (not like "M", "i", "n", "o", ... but like "Min", "or", and "u")

```{r}
input <- "Creative Copy Matsui Minoru" |> 
  str_to_upper()
randomize_N_times <- 100
nwords <- input |>  
  str_split(pattern = "\\s") |> 
  length()
# init out of for
result <- vector(length = randomize_N_times)
input_vec <- input |> 
  str_split(pattern = "\\s") |> 
  unlist()
for(n in 1:randomize_N_times) {
  # initialize
  result_vecs <- vector(length = nwords)
  for(i in 1:length(input_vec)) {
    input_str <- input_vec[i] 
    length_of_str <- nchar(input_str)
    split_letters <-input_str |>
      str_split(pattern = "") |> 
      unlist()
    if (length_of_str > 2) {
    middle_section <- split_letters[2:(length(split_letters) - 1)] |> 
      sample() |> 
      paste(collapse = "")
    } else {middle_section <- ""}
      result_vec <- paste0(split_letters[1], middle_section, split_letters[length(split_letters)])
      # result_vec = "Mrionu"
      result_vecs[i] <- result_vec
  } 
  result_str <- paste(result_vecs, collapse = " ")
  result[n] <- result_str
}
```


# random walk chinese whisperer

change gradually
mutate / combine / insert / 

チェインでどんどんコピーミスが蓄積していく感じをCOPY, COPY, CPOY, CPOPY, CPOPY, CPOPY, CPPY, CPY, CPYP, CPYYP,みたいにやりたい。で、CREATIVE COPY |>  COPY |> CPOPY |> ... としたい。どうやって変異させる？

コードはスライドの最後につけよう。


### mutation rate
```{r}
library(extraDistr) # rbbinom
rbeta(100, ) |> as.tibble() |> 
  ggplot(aes(value)) +
  geom_density()


rbinom(1000, 20, prob = .5) |>
  as.tibble() |> 
  ggplot(aes(value)) +
  geom_histogram(binwidth = 1)

rbbinom(size = 20, n = 100000, alpha = 1.5, beta = 1.5) |> 
  as.tibble() |> 
  ggplot(aes(value)) +
  geom_histogram(binwidth = 1)

runif(100000, min = 0, max = 20) |> ceiling() |> as.tibble() |> 
    ggplot(aes(value)) +
  geom_histogram(binwidth = .1)

rlogis(1000, location = 0, scale = 1) |> as.tibble() |> 
    ggplot(aes(value)) +
  geom_histogram(binwidth = .1)

vec <- rbbinom(size = 20 - 1, n = 100000, alpha = 1.5, beta = 1.5)  + 1
vec |> max()

```

#### sigmoid

[sigmoid at wiki](https://ja.wikipedia.org/wiki/%E3%82%B7%E3%82%B0%E3%83%A2%E3%82%A4%E3%83%89%E9%96%A2%E6%95%B0)

`y_max` mutation_rateは小さめにしたい。1は大きすぎ。
`x_shift` 右にシフトしたい。(x_shift, y_max/2)を中心に点対称になる
`gain` 急さ。1より大きくすると傾きが急になる。

```{r}
sigmoid <- 
  function(x, gain = 1, x_shift = 0, y_max = 1) {
    1 / (1 + exp(gain * (x_shift - x))) * y_max
    } 
base <-
  ggplot() +
  xlim(-3, 15) +
  ylim(0, 1)
base +  geom_function(
  fun = ~ sigmoid(x = .x, gain = .3, x_shift = 5, y_max =.3), size = 2
)  + 
  geom_vline(xintercept = 1) +
  geom_function(
  fun = function(x) {
    (1 / (1 + exp(1.3 * (5 - x)))) * .3
  }, color = "red"
) +
    geom_function(
  fun = function(x) {
    (1 / (1 + exp(-x))) * .3
    }, color = "black"
) +
  geom_function(
  fun = function(x) {
    (1 / (1 + exp(5 - x))) * .3
    }, color = "blue"
)


# sigmoid(x = nchar(trait) - nchar(input), x_shift = nchar(input), gain = 0.5, y_max = .001)
# mu_replace
# sigmoid(
    x = -delta, 
    x_shift = nchar(input), 
    gain = 0.5, 
    y_max = mu_baseline_max)


base +ylim(0, .1) + geom_function(
  fun = ~ sigmoid(x = -.x, gain = 0.5, y_max = .1, x_shift = -10)
) +
  geom_function(
  fun = ~ sigmoid(x = .x, gain = 0.5, y_max = .1, x_shift = 10)
)
```


```{r}
# CONSTANTS
input <- "XERROXCOPY" |> str_to_upper()
timesteps <- 1000
beta_param <- 1.5 # 1.5 seems fine
input_vec <- input |>
  str_split(pattern = "") |> 
  unlist()
result <- vector(length = timesteps)

mu_baseline_max <- .1
trait <- input
runs <- 2 # one for a sequence, and one for reverse sequence
results <- list()
for(run in 1:runs) {
  trait <- input
for(timestep in 1:timesteps) {
  delta <- nchar(trait) - nchar(input)
  # 1. replace just one char with something from pool (input_split)
  mu_replace <- mu_baseline_max
  should_replace_somewhere <- runif(1) < mu_replace
  if (should_replace_somewhere) {
    replace_where <- rbbinom(n = 1, size = nchar(trait) - 1, alpha = beta_param, beta = beta_param) + 1 # + 1 to prevent 0. without it it gives 0 to nchar(trait) if size = nchar(trait).
    replace_with_char <- sample(input_vec, 1)
    trait <- trait |> 
      stri_sub_replace(from = replace_where, length = 1, value = replace_with_char)
  }
  
  # variable mutation strengths: insert and remove
  # insert: add c, o, p, or y in arbitrary position.
  # https://stackoverflow.com/questions/13863599/insert-a-character-at-a-specific-location-in-a-string
  mu_add <- sigmoid(
    x = -1 * delta,
    x_shift = -1 * nchar(input), # when x = -delta, x_shift also must be -
    gain = 0.5,
    y_max = mu_baseline_max) #  should be stronger if length < 4
  # print(paste('mu_add', mu_add))
  should_insert_somewhere <- runif(1) < mu_add
  if (should_insert_somewhere) {
    insert_where <- 
      rbbinom(n = 1, size = nchar(trait) - 1, alpha = beta_param, beta = beta_param) + 1
    trait <- trait |>
      stri_sub_replace(
        from = insert_where,
        to = insert_where - 1,
        value = sample(input_vec, 1))
  }
  # remove
  mu_remove <- sigmoid(delta) # delete the letter. should be stronger if length >> 4
  should_remove_somewhere <- runif(1) < mu_remove
  if (should_remove_somewhere) {
    remove_where <- rbbinom(n = 1, size = nchar(trait) - 1, alpha = beta_param, beta = beta_param) + 1
    trait <- trait |>
      stri_sub_replace(
        from = remove_where,
        length = 1,
        value = ""
      )
  }
  # reverse
  mu_reverse <- .001
  should_reverse <- runif(1) < mu_reverse
  if (should_reverse) {
  trait <- trait |> stri_reverse()

  }
    
  # todo: add random character? 
  # https://rdrr.io/cran/stringi/man/stri_rand_strings.html
  # stri_rand_strings(1, 1, '[A-Z]')
  
  # store the result for output
  
  result[timestep] <- trait
}
  results[[run]] <- result
}
```

### check if ok

```{r}

result[10000]
nchar(input)
result |> 
  nchar() |> 
  as.tibble() |> 
  rowid_to_column(var = "timestep") |> 
  ggplot(aes(timestep, value)) +
  geom_line()
```

### output


```{r}
library(data.table)
result <- results[[1]] |> unlist() |> paste(collapse = " > ")
result_reverse<- results[[2]] |> unlist() |> rev() |> paste(collapse = " > ")
result_str <- paste(result_reverse, result)
fwrite(list(result_str), file = "chinese_whisperer.csv")

```

### output animation

```{r}
result <- results[[1]] |> unlist()
result_reverse<- results[[2]] |> unlist() |> rev()
result_in_on_vec <- c(result_reverse, result)
df <- result_in_on_vec |> 
  as_tibble() |> 
  rowid_to_column(var = "timestep")
```

#### animation


```{r}

animation1 <- df |> 
  ggplot(aes(x = 0, y = 0 )) +
  geom_text(aes(label = value), family = "Fira Code", size = 10) +
  geom_text(aes(x = 0, y = 0.1, label = timestep), family = "Fira Code", size = 3)  +
  xlim(c(-0.5, 0.5)) +
  ylim(c(-0.5, 0.5))  +
  theme_void(base_family = "Fira Code") +
  transition_states(
    timestep,
    transition_length = 0,
    state_length = 0.1
  )
animation1 |> 
  animate(
    fps = 60,
    nframes = 2000,
    # renderer = ffmpeg_renderer()
    renderer = ffmpeg_renderer()
    )
anim_save(filename = "xerroxanimate.mp4")
```



